# Results

Linear regression models were created based on the traffic analysis parameters of occupancy, flow, and density, as seen in Table \@ref(tab:models-summ). This table compares..

```{r models-summ, echo=FALSE}
tar_load(modelsummary)

model_summary <- function(linear_models){
  modelsummary(
    linear_models, 
    estimate = "{estimate} ({statistic}){stars}",
    statistic = NULL, title = "Optim_K Regression Estimates",
    notes = c("t-statistics in parentheses, * p < 0.1, ** p < 0.05, *** p < 0.01"),
    escape = FALSE
  )
}
```

A cluster analysis was also performed, which compares the flow versus the density for each 15-minute period and then compares the average K for each cluster, as seen in Table \@ref(tab:cluster-analysis-tab) and Figure \@ref(fig:cluster-analysis-fig). It was found that in general, the higher the traffic density, the more accurate the conservation model became, in other words, K began to approach 0. As traffic density decreased on the ramp, K began to increase closer to 1. With this information, a heuristic model can be developed to predict the queue length of vehicles at the metered on-ramp based on the traffic density on the ramp.

```{r cluster-analysis-tab}
## still need to clean up table to be presentable
tar_load(cluster_estimates)

estimate_cluster_k <- function(cluster_data){
  cluster_data %>% 
    group_by(cluster) %>%
    summarise(
      mean_k = mean(optim_k), n = n(), sd = sd(optim_k),
      max_density = max(density), min_density = min(density)
    ) %>%
    arrange(mean_k)
}

```

```{r cluster-analysis-fig}
tar_load(cluster_plot)

plot_clusters <- function(cluster_data){
  ggplot(cluster_data, aes(x=density, y = flow, col = factor(cluster))) + 
    geom_point()
}

```


```{r pdata_plot, echo=FALSE}
## still need to figure out how to plot this correctly
library(targets)
library(tidyverse)
library(readxl)
tar_load(plot_data)

ggplot(plot_data %>% mutate(period = str_c(day, hour, period)), 
       aes(x = time, y = queue_length, color = name)) + 
  geom_line() +
  facet_wrap(~period)
```

## Linear Regression

## Cluster Analysis / Heuristics
